Investigating the Impact of Macroeconomic Indicators on the FTSE 250 Index Using Statistical and Machine Learning Methods

### 2024-08-25 tex/paper.tex
structure revised + final minor edits

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
Section 3 presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
the study.

@@

\subsubsection{Machine Learning Methods for Forecasting}

@@

\section{Data}

Monthly FTSE 250 Index price data between Oct 1992 to March 2024 has been utilised for this study giving 379 observations.

@@

For the interest rate data, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
does not convene on a monthly basis, and are not obligated to 
announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using \algoref{interest_rate_adjustment}.

@@

Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
increasing role played by monetary policy in the UK since May 1997 when the Bank of England was given operational responsibility for setting interest rates \citep{king1997changes}. The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
subsequent post-COVID inflationary pressures.

@@

\section{ARDL Model}

Based on the empirical research and characteristics of the UK economy,
the stock market function is defined as:

@@

The study uses an autoregressive distributed lag (ARDL) model in conjunction with 
an error correction model (ECM) to determine 
whether a long-term dependence exists for the FTSE250 and the selected macroeconomic 
indicators, and if so to quantify the long-run and short-run relationships
between the former and the latter.

\subsection{Methodology}

Prior to constructing the ARDL model a natural logarithmic transform was applied
on the variables. This was done so that the coefficients appear as elasticities, 
which can be interpreted easily as percentage changes. Subsequently,
the level at which the variables are stationary was ascertained using the 
Augmented Dickey-Fuller (ADF) unit root test. The optimum lag for the variables in the ARDL model was then determined using the Schwarz
Information Criterion.

@@

\begin{align*}
    \Delta \text{FTSE}_{250_t} &= \alpha_0 + \sum_{i=1}^{p} \alpha_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \beta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \beta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \varepsilon_t + \phi_{1} FTSE_{250_{t-1}} \\
                               & + \phi_{2} CPI_{t-1} + \phi_{3} INT_{t-1} +\phi_4 EXCHG_{t-1} + \phi_5 M3_{t-1}
\end{align*}

Here, the intercept term was given by $\alpha_0$. The 
remaining $\alpha_i$ terms represent the short-run relationships 
between the variables, and $\phi_j$ represent the long-run relationships.

In the ARDL bounds test, the null and alternate hypothesis were:

\begin{align*}
    H_{0}: \phi_i &= \phi_j\\
    H_{1}: \phi_i &\neq \phi_j
\end{align*}

for, $i,j = 1,2,3,4,5$, with $i\neq j$. The F-statistic was computed for the joint
significance of the lagged variables and compared against the critical values
given by \citeboth{pesaran2001} to determine whether cointegration exists, 
and at what significance level. If cointegration exists, then 
the long-run coefficients $\kappa_i$ are estimated to find the cointegration equation:

\begin{align}
    FTSE_{250} = \kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon \label{eq:two}
\end{align}

Then, an Error Correction Model (ECM) was constructed:

\begin{align*}
    \Delta \text{FTSE}_{250_t} &= \alpha_0 + \sum_{i=1}^{p} \alpha_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \beta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \beta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \gamma\text{ECT}_{t-1} + \varepsilon_t
\end{align*}

where the error correction term, ECT, was found by calculating the residual 
produced by \eqref{eq:two}:

\begin{align*}
    ECT_{t-1} = FTSE_{250_{t-1}} - \biggl(\kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon \biggr)
\end{align*}

The parameter $\gamma$ describes the speed at which short-run deviations from the long-run equilibrium
are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that converges the deviation 
to the long-run equilibrium. To estimate the short-run coefficients, $\alpha_i$ and $\beta_k$, as 
well as the short-run correction speed $\gamma$ Ordinary Least Squares (OLS) regression was used.

Finally, diagnostic tests are conducted to confirm that the model does not
display autocorrelation, heteroskedascity, non-normality and that the cumulative 
sum of residuals and squared residuals are both stable.

\subsection{Results}

The null hypothesis of the ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
is that the variables are stationary.

\begin{table}[h!]
    \centering
    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
    at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}}
    \begin{tabular}{lccc}
        \toprule
        \textbf{Variable} & \textbf{Level ADF} & \textbf{First Difference ADF} & \textbf{Stationarity} \\
        \midrule
        FTSE$_{250}$ & $-3.2646^{****}$ & $-6.6892^{*}$ & At first difference \\
        CPI          & $-1.4594$ & $-4.705^{*}$ & At first difference \\
        EXCHG        & $-2.2467$ & $-7.3855^{*}$ & At first difference \\
        INT          & $-2.0946$ & $-4.1156^{*}$ & At first difference \\
        M3           & $-0.70827$ & $-6.3806^{*}$ & At first difference \\
        \bottomrule
    \end{tabular}
\end{table}

Subsequently, the optimum lag of the variables was calculated using the 
Schwarz Information Criterion, to obtain the optimum model 
of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.1$ was compared against the critical values
given by \citeboth{pesaran2001}, providing evidence of cointegration 
at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
level.

The long-run and short-run coefficients were then estimated, and presented 
alonsgide the diagnostic test results which confirm that the model does not
display autocorrelation, heteroskedascity, non-normality and the cumulative 
sum of residuals and squared residuals are both stable. The error
correction term is also given, and it was negative, as expected.

The long-term coefficients show that the interest rate and exchange rate
interest rate have significant long-term effects on the stock market price
while money supply and inflation are statistically insignificant. \citeboth{neifar2023}
proposes that this is because deflation, that is the negative movements in the 
CPI, is a statistically significant factor in the movement of the UK stock 
market, but positive movements (inflation) are not. Nevertheless, the positive
coefficient of 3.16, and the negative short-term adjustment, is in line with the expectation from the literature that 
inflation exerts a positive effect on the stock market in the long-term, but a negative 
one in the short-run.

The insignificant long-term effect of the money supply can be explained by several reasons.
Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
money supply, at least over certain intervals if not always. This would mean anticipated changes in the money supply would not affect the stock market price, only the
unanticipated component of a change in money supply would, as established by \citeboth{sorensen1982} and \citeboth{bernanke2005}. 
It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
on the stock market dynamics. In the short-run it is observed that the money supply does have a 
significantly negative effect on the stock market, as outlined by \citeboth{olawale2014}. According to \citeboth{sellin2001}, 
this supports the Keynesian argument that a positive money supply shock will lead people to
expect a contraction in future monetary policy. 
This leads to an increase in the demand for funds, which drives up the current rate of interest, in turn
reducing the present value of earnings in the Discounted Cash Flow (DCF) stock valuation model. The effect is a decline in the stock market prices.

Unsurprisingly, the USD/GBP exchange rate was significantly related to the FTSE250's movements. It 
should not be forgotten that, although it is considered to be a more domestic indicator of the LSE, 
at times $50\%$ of the FTSE250's earnings have come from abroad, indicating that 
the value of the Pound has a significant impact on this stock market. 
In the long run, the appreciation of the Pound would increase the relative 
cost of UK exports increases, thereby reducing the 
competitiveness of FTSE250 companies in international markets. This in turn would lower their earnings
and depress the FTSE250. Moreover, foreign investors may view the stronger Pound as a signal of reduced returns on their investments in the UK, which can lead to decreased investment in the FTSE250. 
In the short-term, however, before structural shifts can materialise the
weakening of the Pound has a positive effect. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it will lead to increased revenue in Pound terms for FTSE250 companies with substantial overseas operations. This is because their foreign earnings, when converted back into Pounds, will be worth more, potentially increasing profits. The findings 
are in line with \citeboth{neifar2023} and \citeboth{khan2018}.

\section{RBF Neural Network}

\subsection{Model Architecture}

@@

The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
a $c$-dimensional target space $\boldsymbol{\mathrm{y}}$. The dataset consists of $N$ input
vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
values $y_n$. The objective is to find a function
  $h(\boldsymbol{\mathrm{x}})$ such that:

\subsection{Methodology}

The radial basis function neural network was utilised for both short-term and long-term 
forecasting. For the short-term, 
the model was tested over a 5-month horizon, based on the optimum lag calculated earlier. 
The long-term forecst spanned 57 months. 
The data was differenced based on the recommendations of the 
unit root tests conducted earlier to achieve stationarity.

The data was normalised using a z-score normalisation, because neural networks work best with data
that is between a specified range (\citeboth{mendelsohn1993}). Following this 
it was necessary to program the radial basis function neural network into Python. The kernel regression approach outlined by 
\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network.

\subsubsection{The Gaussian Process Regressor}

@@

where $y$ differs from the function values $f(x)$ by noise which is assumed
to follow an iid Gaussian distribution (\citeboth{rasmussen2006}). 
$\mathcal{N}_1$
\begin{align*}
    \varepsilon \sim \mathcal{N}_1 \bigl(0, \sigma_{n}^2\bigr)
\end{align*}
This assumption together with the model produces the 
probability density of the observations given the inputs, 
which is factored over cases in the training set. This is
permissible because of the independence assumption (\citeboth{rasmussen2006}).

@@

\subsubsection{The Network}

First, the optimal hyperparameters in the covariance function \eqref{eq: gimq} 
are computed using GridSearch Cross Validation; these hyperparameters are:
\begin{enumerate}
    \item $\alpha$: a regularisation parameter representing a small quantity
    that is added to the diagonals of the covariance matrix to ensure it is 
    invertible. 
    \item $\beta$: the power in the covariance function, which controls the shape of the function.
    \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for. It moderates the risk of overfitting and underfitting.
    \item Length scale: a parameter that controls the influence that each kernel is to have. Analogous to 
    the width each radial basis function neuron possesses.
\end{enumerate}

Then, training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) are successively passed through the 
covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
all training observations, $K$.

A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
calculates the posterior weights distribution. 
Each possible weight vector $w$ represents a different function, explicitly 
based on the radial basis function covariance function, that could fit the data. In the 
hidden layer the expectation is then taken, which
combines all weights and finds the mean, weighted by the 
respective probabilities. This represents the best functional estimate of the price movement. Based on this estimate 
the predicted output generated.

The testing process is conducted similarly. The 
posterior weight distribution that contains the information of the training phase becomes the new 
prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
are calculated, and then the predictive output is found (\citeboth{rasmussen2006}).

\begin{align}
    p(y_{*}|x_{*}, X, y) = \mathcal{N}(\sigma_{n}^{2} x_{*}^{T}A^{-1}Xy, x_{*}^{T}A^{-1}x_{*}) \label{eq: testw}
\end{align}

\subsection{Results}

The model returned a mean absolute error of MAE $=0.16$ and 
$R^2 = 9\%$ for the long-term prediction, providing a 
$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
in Figures \ref{fig:longmonthly} and \ref{fig:longdaily}, against monthly and daily FTSE250 changes respectively.

\begin{figure}[H]
    \centering
    \input{long_term_monthly.tex}  
    \caption{Long-term predicted vs real monthly FTSE250 changes}
    \label{fig:longmonthly}
\end{figure}

\begin{figure}[H]
    \centering
    \input{long_term_daily.tex}  
    \caption{Long-term predicted vs real daily FTSE250 changes}
    \label{fig:longdaily}
\end{figure}

The short-term prediction yielded MAE = $0.07$ and 
$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figures \ref{fig:shortmonthly} 
and \ref{fig:shortdaily}, against monthly and daily FTSE250 changes respectively.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possess
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price in the short-run.

\begin{figure}[H]
    \centering
    \input{short_term_monthly.tex}  
    \caption{Short-term predicted vs real monthly FTSE250 changes}
    \label{fig:shortmonthly}
\end{figure}

\begin{figure}[h!]
    \centering
    \input{short_term_daily.tex}  
    \caption{Short-term predicted vs real daily FTSE250 changes}
    \label{fig:shortdaily}
\end{figure}

In the short-term the market movements
are more closely tied to recent deviations in the economy. This is because,
theoretically, 
the short-run is not exposed to the various unforeseen external shocks and structural changes
that dilute the relationship between these variables and the stock market in the 
long-run. As the prediction window expands, 
more uncertainty and unknowns are introduced, which translates into 
lower predictive quality. Furthermore, as the time horizon extends into the long-run, fundamental factors 
begin to matter a lot more. The compound effect of earnings, cash flows,
innovations and managerial efficiencies begins to outweigh the 
influence of fluctuating macroeconomic variables. Essentially, firms become accustomed 
to the changed economic environment and over time. Thus, in the short-run 
a different phenomenon is observed.

>>>>

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
Section \ref{sec:meth} presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
the study.

@@

\subsubsection{Machine Learning Methods for Forecasting}
\label{subsubsec: fore}

@@

\section{Methodology}
\label{sec:meth}

\subsection{Data}
Monthly FTSE 250 Index price data between Oct 1992 to March 2024 has been utilised for this study giving us a dataset with 379 observations.

@@

For the interest rate data, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
does not convene on a monthly basis, and are not obligated to 
announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using \ref{alg:interest_rate_adjustment}.

@@

Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
increasing role played by monetary policy in the UK since May 1997 when the Bank of England was given operational responsibility for setting interest rates (\citeboth{king1997changes}). The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
post-COVID inflationary pressures.

@@

\subsection{ARDL-ECM Model}

The study uses an autoregressive distributed lag (ARDL) model in conjunction with 
an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
and to quantify these relationships. Prior to constructing the model a natural logarithmic transform was applied
on the variables. This was done so that the coefficients can be interpreted easily as percentage changes.

Based on the selected macroeconomic indicators,
the stock market function was defined as:

@@

The ARDL model specifies the dependent variable as a function of its own 
lagged values and the lagged values of independent variables. 
The variables are permitted to be of differing stationarity. The stationarity 
of the variables was tested using the Augmented Dickey-Fuller 
Unit Root Test found within the tseries package in R. The optimal lags for each of the 
variables were calculated using the Schwarz Information Criterion, contained
within the auto\_ardl function in the ARDL package in R.

@@

\begin{align*}
    \Delta \text{FTSE}_{250_t} &= \alpha_0 + \sum_{i=1}^{p} \alpha_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \beta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \beta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \varepsilon_t 
\end{align*}

The intercept term was given by $\alpha_0$. The 
remaining $\alpha_i, \beta_i$ coefficients represent the long-run relationships 
between the variables. The coefficients were computed using the ardl function in the 
ARDL package in R.

The short-term dynamics were modelled using an Error Correction Model (ECM)

\begin{align*}
    \Delta \text{FTSE}_{250_t} &= \gamma_0 + \sum_{i=1}^{p} \gamma_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \delta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \delta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
                               & + \sum_{j_{3}=0}^{q_3} \delta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \delta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \lambda \text{ECM}_{t-1} + \varepsilon_t
\end{align*}

Here, the intercept term was given by $\gamma_0$. The 
remaining $\gamma_i, \delta_i$ coefficients represent the short-run 
relationships between the variables. The $\lambda$ coefficient 
describes the speed at which short-run deviations from the long-run equilibrium
are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that guides the deviation 
to the long-run equilibrium. These coefficients, as well 
as the error correction term ECT, were computed using the recm function in the 
ARDL package in R.

Subsequently, cointegration was checked. The null and alternate hypothesis were:

\begin{align*}
    H_{0}: \beta_i &= 0\\
    H_{1}: \beta_i &\neq 0
\end{align*}

for all $i = 1,2,3,4$. The bounds\_f\_test in 
the ARDL library, based on the bounds test approach 
outlined in \citeboth{pesaran2001}, was used to calculate the 
F-statistic of the ARDL model and subsequently test for cointegration.

Finally, diagnostic tests are conducted to confirm that the model does not
display autocorrelation, heteroskedascity, non-normality and that the cumulative 
sum of residuals and squared residuals are both stable. This is to 
ensure our regression is not spurious.

\subsection{RBF Neural Network}

Radial basis function neural networks are a type of artificial 
neural network that use radial basis functions at each node in the 
activation layer of the network. They are able to identify complex, 
non-linear relationships between time-series variables and have been used
extensively in this aspect (\ref{subsubsec: fore}). The advantages of radial basis function neural networks over other 
neural networks in time-series forecasting lies in their 
ability to operate effectively on small sample sizes.

\subsubsection{Model Architecture}
The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
a $c$-dimensional target space $\boldsymbol{\mathrm{y}}$. The dataset consists of $N$ input
vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
values $y_n$. The objective is to find a function
  $h(\boldsymbol{\mathrm{x}})$ such that:

@@

\subsubsection{Methodology}

First, the dataset was normalised using a z-score normalisation, because neural networks work best with data
that is between a specified range (\citeboth{mendelsohn1993}).

The inputs into the network are 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\),

\begin{align}
    \boldsymbol{\mathrm{x_i}} = \{\Delta \text{CPI}_i, \Delta \text{EXCHG}_i, \Delta \text{INT}_i, \Delta \text{M3}_i\}
\end{align}

where \(\Delta \text{CPI}_i\), \(\Delta \text{EXCHG}_i\), \(\Delta \text{INT}_i\), and \(\Delta \text{M3}_i\) 
represent differenced monthly values in the CPI, USD/GBP exchange rate, interest rate, and M3 money supply, respectively.

The output is a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\), representing differenced monthly values in the FTSE250.

\begin{align}
    \boldsymbol{\mathrm{y_i}} = \{\Delta \text{FTSE}_{250_i}\}
\end{align}

Following this 
it was necessary to program the radial basis function neural network into Python. The Bayesian kernel regression approach outlined by 
\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network. For a 
detailed description of this construction please refer to the \hyperref[sec: appendix]{Appendix}.

Based on this, the train-test process is then conducted:
\begin{enumerate}
    \item First, the optimal hyperparameters in the covariance function \eqref{eq: gimq} 
    are computed using GridSearch Cross Validation; these hyperparameters are:
    \begin{enumerate}
        \item $\alpha$: a regularisation parameter representing a small quantity
        that is added to the diagonals of the covariance matrix to ensure it is 
        invertible. 
        \item $\beta$: the power in the covariance function, which controls the shape of the function.
        \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for. It moderates the risk of overfitting and underfitting.
        \item Length scale: a parameter that controls the influence that each kernel is to have. Analogous to 
        the width each radial basis function neuron possesses.
    \end{enumerate}
    \item Then, training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) are successively passed through the 
    covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
    all training observations, $K$.
    \item A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
    calculates the posterior weights distribution. 
    Each possible weight vector $w$ represents a different function, explicitly 
    based on the radial basis function covariance function, that could fit the data. In the 
    hidden layer the expectation is then taken, which
    combines all weights and finds the mean, weighted by the 
    respective probabilities. This represents the best functional estimate of the price movement. Based on this estimate 
    the predicted output generated.
    \item The testing process is conducted similarly. The 
    posterior weight distribution that contains the information of the training phase becomes the new 
    prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
    are calculated, and then the predictive output is found (\citeboth{rasmussen2006}).

    \begin{align}
        p(y_{*}|x_{*}, X, y) = \mathcal{N}(\sigma_{n}^{2} x_{*}^{T}A^{-1}Xy, x_{*}^{T}A^{-1}x_{*}) \label{eq: testw}
    \end{align}
\end{enumerate}

A short-term and long-term forecast is conducted using this process. For the short-term forecast a test set of 5 months was used, and for the long-term forecast a test set of 57 months was used.

\section{Results}

\subsection{ARDL-ECM}

\begin{table}[h!]
    \centering
    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in natural logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
    at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}}
    \begin{tabular}{lccc}
        \toprule
        \textbf{Variable} & \textbf{Level ADF} & \textbf{First Difference ADF} & \textbf{Stationarity} \\
        \midrule
        FTSE$_{250}$ & $-3.2646^{****}$ & $-6.6892^{*}$ & At first difference \\
        CPI          & $-1.4594$ & $-4.705^{*}$ & At first difference \\
        EXCHG        & $-2.2467$ & $-7.3855^{*}$ & At first difference \\
        INT          & $-2.0946$ & $-4.1156^{*}$ & At first difference \\
        M3           & $-0.70827$ & $-6.3806^{*}$ & At first difference \\
        \bottomrule
    \end{tabular}
\end{table}

The optimum lags for the ARDL model were ($FTSE_{250} = 5$, $CPI = 4$, $EXCHG=7$, $INT =7$, $M3=4$) for  The resulting F-statistic of $4.1$ was compared against the critical values
given by \citeboth{pesaran2001}, providing evidence of cointegration 
at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
level.

\subsubsection{Short-term Impacts}

The interest rate has a significantly
negative short-term impact on the FTSE250 price in the short term. 
This can be attributed to the immediate effect of interest rate hikes, 
which decrease the present value of future earnings according to the 
Discounted Cash Flow (DCF) valuation model, thereby 
affecting investor sentiment and depressing stock prices. 
This finding is consistent with the studies by 
\citeboth{alam2009}, \citeboth{demir2019}, and \citeboth{neifar2023}.

The money supply has a significantly negative short-term effect on the 
FTSE250 price, as noted by \citeboth{olawale2014}. 
According to \citeboth{sellin2001}, this supports the Keynesian argument 
that an increase in money supply leads to expectations of a future 
monetary contraction. This expectation increases the demand for 
funds, driving up current interest rates and reducing stock prices 
through the same mechanism as described previously.

In the short-term, before structural shifts can materialise, the
weakening of the Pound against the Dollar has a positive, but 
insignificant effect. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it will lead to increased revenue in Pound terms for FTSE250 
companies with substantial overseas operations. 
This is because their foreign earnings, 
when converted back into Pounds, will be worth more, 
potentially increasing profits. However, the insignificant impact suggests 
that FTSE250 firms are resilient to short-term fluctuations in the stock market 
price. The possible reasons for this are that they have diversified 
operations, meaning fluctuations against the Dollar will not threaten 
their profits and they use hedging strategies involving financial instruments 
such as currency forwards which protect against volatility.

The insignificance of inflation, defined as a positive 
increase in the CPI, confirms the findings of 
\citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
The CPI as a measure does not uniformly represent all sectors of the economy, 
and may consequently be unrepresentative of inflationary changes in the 
sectors that predominate the FTSE250. In addition, different sectors 
are more vulnerable to inflation than others which may distort the effect 
of inflation even further. Moreover, inflation may be anticipated 
by investors and already be factored into stock prices.

\subsubsection{Long-term Impacts}

The CPI did not have a significant effect on the FTSE250 price over the 
long-run, lending support to the findings of 
\citeboth{gultekin1983} and \citeboth{firth1979}
that the Fisher hypothesis is valid for the UK. One of the 
reasons for this may be that, since stock prices represent the 
discounted value of future cash flows (which are real), 
they are not directly impacted by inflation in the long run. Moreover, 
firms are able to pass on nominal costs to consumers which reduces
the impact of inflation.
The freer role of the Bank of England in setting monetary policy could be 
another factor in this, as a stricter monetary policy would keep inflation
in check and obscure any long-term equilibrium relationship.

The insignificant long-term effect of the money supply can be explained by several reasons.
Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
money supply, at least over certain intervals if not always. This would mean anticipated changes in the money supply would not affect the stock market price, only the
unanticipated component of a change in money supply would, as established by \citeboth{sorensen1982} and \citeboth{bernanke2005}. 
It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
on the stock market dynamics.

The USD/GBP exchange rate significantly positively impacted the FTSE250 price, confirming the findings of 
\citeboth{wong2022}. That there is a 
significant relationship is not surprising, because whilst the FTSE250 is a more domestic indicator of the London Stock Exchange than the FTSE 100 Index, 
at times nearly $50\%$ of its earnings have come from abroad, indicating that 
the value of the Pound has a significant impact on this stock market. There are many reasons for why this effect is observed.
In the long run, the appreciation of the Pound would decrease the 
cost of imports, thereby increasing the 
profits of FTSE250 companies and inflating their share prices.
Moreover, a stronger Pound makes UK assets more appealing to foreign investors, as their investments would be worth more when converted back to their domestic currencies.
This would drive up demand for UK stocks and positively influence the 
FTSE250 price.

\subsection{RBFNN}

\subsubsection{Long-term Forecast}
The model returned a mean absolute error of MAE $=0.16$ and 
$R^2 = 9\%$ for the long-term prediction, providing a 
$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
in Figure \ref{fig:longmonthly}.

\begin{figure}[H]
    \centering
    \hspace*{-1.5cm}
    \input{long_term_monthly.tex}  
    \caption{Long-term predicted vs real monthly FTSE250 changes}
    \label{fig:longmonthly}
\end{figure}

\subsubsection{Short-term Forecast}

The short-term prediction yielded MAE = $0.07$ and 
$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figure \ref{fig:shortmonthly}.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possess
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price in the short-run.

\begin{figure}[H]
    \centering
    \hspace*{-1.5cm}
    \input{short_term_monthly.tex}  
    \caption{Short-term predicted vs real monthly FTSE250 changes}
    \label{fig:shortmonthly}
\end{figure}

In the short-term the market movements
are more closely tied to recent deviations in the economy. This is because,
theoretically, 
the short-run is not exposed to the various unforeseen external shocks and structural changes
that dilute the relationship between these variables and the stock market in the 
long-run. As the prediction window expands in regression problems, 
more uncertainty and unknowns are introduced, which translates into 
lower predictive quality. Furthermore, as the time horizon extends into the long-run, fundamental factors 
begin to matter a lot more. The compound effect of earnings, cash flows,
innovations and managerial efficiencies begins to outweigh the 
influence of fluctuating macroeconomic variables. Essentially, firms become accustomed 
to the changed economic environment and over time. Thus, in the short-run 
a different phenomenon is observed.

\section{Appendix}
\label{sec: appendix}

@@

where $y$ differs from the function values $f(x)$ by noise which is assumed
to follow an iid Gaussian distribution 
$\mathcal{N}_1$ (\citeboth{rasmussen2006}).
\begin{align*}
    \varepsilon \sim \mathcal{N}_1 \bigl(0, \sigma_{n}^2\bigr)
\end{align*}
This assumption together with the model produces the 
probability density of the observations given the inputs, 
which is factored over cases in the training set. This is
permissible because of the independence assumption (\citeboth{rasmussen2006}).

@@
